---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
```{r}
library(haven)
library(ggplot2)
library(rstatix)
library(tidyverse)
library(skimr)
library(descr)
library(labelled)
url <- "https://github.com/drCES/winter_603/raw/main/California%20Real%20Estate%20Data%20-%202015%20Ahmed.dta"

# Import the data into R
data <- read_dta(url)
```

# Setup
```{r}
head(data)
tail(data)

summary(data)
skim(data)
```

```{r}
freq(data$bedrooms_5)
freq(data$bathrooms_rnd)
```

```{r}
ggplot(data, aes(x = sqft)) +
  geom_histogram(aes(y = after_stat(density)), position = "identity", bins = 10, alpha = 0.25, color="black") +
  labs(title = "Histogram Plot of House Square Ft",
       x = "Square Ft",
       y = "Density") +
  theme_minimal()

ggplot(data, aes(x = price)) +
  geom_histogram(aes(y = after_stat(density)), position = "identity", bins = 10, alpha = 0.25, color="black") +
  labs(title = "Histogram Plot of House Price",
       x = "Price",
       y = "Density") +
  theme_minimal()
```

# Part 1: Descriptive Data Review:


1) Describe what type of variable each is (continuous, nominal, ordinal, etc)
- sqft: continuous. ratio scale
- price: continuous. ratio scale
- bedroom_5: discrete & ordinal. ratio scale
- bathrooms_rnd: discrete & ordinal. ratio scale
2) Describe what type of probability function is appropriate for that variable (PMF or PDF)
- sqft: PDF
- price: PDF
- bedroom_5: PMF
- bathrooms_rnd: PMF
3) Create a histogram that shows the distribution of each variable


```{r}
ggplot(data, aes(x = bathrooms_rnd)) +
  geom_histogram(aes(y = after_stat(density)), position = "identity", alpha = 0.25, color="black") +
  labs(title = "Histogram Plot of House Bathrooms",
       x = "Bathrooms",
       y = "Density") +
  theme_minimal()

ggplot(data, aes(x = bedrooms_5)) +
  geom_histogram(aes(y = after_stat(density)), position = "identity",  alpha = 0.25, color="black") +
  labs(title = "Histogram Plot of House bathrooms_rnd",
       x = "Bedrooms",
       y = "Density") +
  theme_minimal()

ggplot(data, aes(x = sqft)) +
  geom_histogram(aes(y = after_stat(density)), position = "identity", bins = 30, alpha = 0.25, color="black") +
  labs(title = "Histogram Plot of House Square Ft",
       x = "Square Ft",
       y = "Density") +
  theme_minimal()

ggplot(data, aes(x = price)) +
  geom_histogram(aes(y = after_stat(density)), position = "identity", bins = 30, alpha = 0.25, color="black") +
  labs(title = "Histogram Plot of House Price",
       x = "Price",
       y = "Density") +
  theme_minimal()
```
4) Calculate the mean and standard deviation of each variable

```{r}
skim(data)
sprintf("Mean of sqft: %.2f", 2364.9046728972)
sprintf("Standard Deviation of sqft: %.2f", 1224.55698183349)

sprintf("Mean of price: %.2f", 589362.811214953)
sprintf("Standard Deviation of price: %.2f", 509026.067649057)

sprintf("Mean of bedroom: %f", 3.36635514018692)
sprintf("Standard Deviation of bedroom: %.2f", 1.07420765375093)

sprintf("Mean of bathroom: %f", 2.79065420560748)
sprintf("Standard Deviation of bathroom: %.2f", 1.05694889043006)
```

5) Find the modal number of bedrooms and bathrooms in the dataset

- bedroom: 3 bedrooms is the most frequent
- bathroom: 2 bathrooms is the most frequent

```{r}
freq(data$bedrooms_5, plot=F)
freq(data$bathrooms_rnd, plot=F)
```

6) For the two continuous variables in the dataset,
- Create a 95% confidence interval (Decide if you will use the Gaussian or Student's T distribution first)
```{r}
# price
ci_price <- t.test(data$price, conf.level = 0.95)
print("95% confidence interval for House Price")
ci_price$conf.int

ci_sqft <- t.test(data$sqft, conf.level = 0.95)
print("95% confidence interval for House Square Ft")
ci_sqft$conf.int
```

- Calculate the z-score for each individual value and save it as a new column in the dataset
```{r}

data_update <-data %>%
  mutate(
         zscore_price = (price - mean(price))/sd(price),
         zscore_sqft = (sqft - mean(sqft))/sd(sqft))

print(data_update)
```

8) Describe what you see in the 95% confidence intervals
- Price: The 95% confidence interval tells us that there is 95% chance that the true population mean of house price is within the range [546131.7, 632593.9] based on our sample data.
- Price: The 95% confidence interval tells us that there is 95% chance that the true population mean of house square ft is within the range [2260.904, 2468.905] based on our sample data.

#Part 2: Hypothesis Creation

1) H0 = Average Price of Home is the Same for Houses with 3 Bedrooms vs 4 Bedrooms, HA = Average Price of Home is NOT the Same for Houses with 3 Bedrooms vs 4 Bedrooms
   - HA = Average Price of Home with 4 Bedrooms > Home with 3 Bedrooms
2) H0 = Average Price of Home is the Same for Houses with Square Footage >= 2364.90 vs Square Footage < 2364.90, HA = Average Price of Home is NOT the Same for Houses with Houses Square Footage >= 2364.90 vs Square Footage < 2364.90
   - HA = Average price of Homes is Greater for Homes with Square Footage >= 2364.90
3) H0 = Average Price of Home is the Same for Houses with 3 Bathrooms vs 4 Bathrooms, HA = Average Price of Home is NOT the Same for Houses with 3 Bathrooms vs 4 Bathrooms
   - HA = Average Price of Home with 4 Bathrooms > Home with 3 Bathrooms
