---
title: "Assignment 2"
output:
  html_document:
    df_print: paged
---

```{r}
library(haven)
library(rstatix)
library(tidyverse)
library(ggplot2)
library(descr)

url <-"https://github.com/drCES/winter_603/raw/main/California%20Real%20Estate%20Data%20-%202015%20Ahmed.dta"

# Import the data into R
data <- read_dta(url)

data_subset <-data %>%
  filter(bedrooms==4 | bedrooms==3)
```

# Part 1
## A. T-Tests:

For 3 and 4 bedroom homes only (meaning create a new subset of the overall data that only includes homes with 3 or 4 bedrooms), run a t-test using the Welch's test looking at differences in...
- Price
- Square footage
- Bathrooms

```{r}
var.test(data_subset$price~data_subset$bedrooms)
var.test(data_subset$sqft~data_subset$bedrooms)
var.test(data_subset$bathrooms~data_subset$bedrooms)
```

```{r}
price_welch <- t.test(data_subset$price ~ data_subset$bedrooms, var.equal = FALSE)
sqft_welch <- t.test(data_subset$sqft ~ data_subset$bedrooms, var.equal = FALSE)
bathrooms_welch <- t.test(data_subset$bathrooms ~ data_subset$bedrooms, var.equal = FALSE)
print(price_welch)
print(sqft_welch)
print(bathrooms_welch)
```


Interpret the results including...
and
Identify which tests include a statistically significant difference between 3 and 4 bedroom homes and explain why that result should be considered statistically significant.

Answer:
- The means for each IV by bedroom
  - Price:
    - The difference in mean Price between 3 Bedrooms and 4 Bedrooms is (585197.2 - 755777.0 = -170579.8)
    - The CI states were are 95% confident the true difference in mean is within the range [-281924.73, -59234.92]
  - Square footage:
    - - The difference in mean Square footage between 3 Bedrooms and 4 Bedrooms is (2020.022 - 2930.690 = -910.668)
    - The CI states were are 95% confident the true difference in mean is within the range [-1141.2412, -680.0942]
  - Bathrooms:
    - - The difference in mean Bathrooms between 3 Bedrooms and 4 Bedrooms is (2.365385 - 3.048276 = -0.682891)
    - The CI states were are 95% confident the true difference in mean is within the range [-0.8421944, -0.5235881]

- The amount of variability there is around each mean
  - Price: 95% confident the true ratio of variance for Price between 3 Bedrooms and 4 Bedrooms is in the range [0.3160191, 0.5881974]
  - Square footage: 95% confident the true ratio of variance for Square footage between 3 Bedrooms and 4 Bedrooms is in the range [0.2393000, 0.4454024]
  - Bathrooms:95% confident the true ratio of variance for Bathrooms between 3 Bedrooms and 4 Bedrooms is in the range [0.4358665, 0.8112661]

- T-test statistic
  - Price: t = -3.018 --> Houses with 4 bedrooms have greater Price (on average) than houses with 3 bedroom
  - Square footage: t = -7.7844  --> Houses with 4 bedrooms have greater Square footage (on average) than houses with 3 bedroom
  - Bathrooms: t = -8.4403  --> Houses with 4 bedrooms have greater Bathrooms (on average) than houses with 3 bedroom

- P-value
  - Price: P-Value 0.002821 < 0.01 --> test is significant. We reject the null that Houses with 3 Bedrooms have same Price as Houses with 4 Bedrooms
  - Square footage: P-Value 2.826e-13 < 0.01 --> test is significant. We reject the null that Houses with 3 Bedrooms have same Square footage as Houses with 4 Bedrooms
  - Bathrooms: P-Value 2.073e-15 < 0.01 --> test is significant. We reject the null that Houses with 3 Bedrooms have same Bathrooms as Houses with 4 Bedrooms

## B. ANOVA:

- Using the full dataset, run an ANOVA with the Bonferroni and TukeyHSD corrections, using bedrooms - specifically the 'bedrooms_5' variable - as the IV and price as the DV.


```{r}
means<-data %>%
  group_by(bedrooms_5) %>%
  summarise(mean = mean(price),
            sd = sd(price),
            n=n())
print(means)

anova_result <- aov(price ~ as.factor(bedrooms_5), data = data)
print(summary(anova_result))

print("====================== TukeyHSD")
TukeyHSD(anova_result)

print("====================== Bonferroni")
pairwise.t.test(data$price, data$bedrooms_5, p.adjust.method = "bonferroni")
```

- Fully interpret what you see in each analysis including:
  - Mean of price by number of bedrooms
    - Bedrooms 1:236529.2109375
    - Bedrooms 2:585197.214285714
    - Bedrooms 3:755777.04137931
    - Bedrooms 4:797268.508196721
    - Bedrooms 5:1068759.05263158
  - Significant differences between the pairwise comparisons
    - Our ANOVA p-value < 2e-16 which is less than 0.01 thus our test is significant. We reject the null that there are no difference in mean between our groups.
      - TukeyHSD:
            - Bedroom groups 3-2, 4-2, 5-2, 6-2, 4-3, 6-3 all have p-vals less than 0.01. These beds our test shows high significance. Indicating their means are not the same
               -   5-3 and 6-4 have vals less than 0.05
      - Bonferroni:
        - 3-2, 4-2, 5-2, 6-2, 2-3, and 6-3 all have p-vals less than 0.01
  - Are there differences in the p-values between the two approaches? If so, which test is more conservative (i.e. which test reports larger p-values)?
    - Yes there are some differences between the tests. It seems the Bonferroni test is slightly more conservative as it has larger p-values
- Which size house, on average, sells for the most money? Explain what you see in the results that allows you to draw that conclusion.
I believe house sizes with 6 bedrooms sell for the most on average. I am to this conclusion because looking at the TukeyHSD results, you see that the differences with houses with 6 bedrooms vs 2,3,4,5 bedrooms is the largest.
# Part 2: Correlations
## Task A:
Using the entire real-estate data (excluding zip code), calculate each of the following correlation types for each variable:
- Pearson’s R
- Spearman’s Rank
- Kendall’s Tau


```{r}
data_new <-data %>%
  dplyr::select(price, sqft, bedrooms, bedrooms_5, bathrooms, bathrooms_rnd)
cor(data_new, method="pearson", use="complete.obs")
cor(data_new, method="spearman", use="complete.obs")
cor(data_new, method="kendall", use="complete.obs")
```

Describe what you see in the correlations. Are some variables correlated at higher rates than others? What do the actual correlation values tell you about the relationships?
- It looks like Price is not highly correlated with any singular variable. All less than .5
- Bedrooms and Bedrooms_5 and Bathrooms and Bathrooms_rnd are very highly correlated with each other. This makes sense because I believe they are slight modifications of the original column
- It seems Square footage is highly correlated with bedrooms and bathrooms. This makes sense as bigger homes can have more bedrooms and bathrooms.
- The correlation values indicate how closely the 2 variables are linearly related to each other. It does not indicate cause and effect.
## Task B:

1) Identify the two “continuous” variables, and decide which is the DV and which is the IV, in the dataset then create a scatter plot putting the appropriate variable on the y and x-axes.
Bonus: Add the correlation coefficient to the plot using the annotate command we reviewed


```{r}
ggplot(data_subset, aes(x = data_subset$sqft, y = data_subset$price)) +
  geom_point(color = "blue", size = 3) +
  labs(title = "Scatterplot Square Footage vs Price",
       x = "Square Footage",
       y = "Price") +
  theme_minimal() + annotate("text",
             x = Inf, y = -Inf,
             label = paste("Pearson's r =", round(cor(x=data_subset$sqft, y=data_subset$price, method="pearson", use="complete.obs"), 2)),
             hjust = 1.1, vjust = -0.5,
             size = 4, color = "red")
```

2) Create a scatter plot between price, as the DV, and bathrooms as the IV. What do you see in this plot? Is it as useful as the plot created in Part 1?

```{r}
ggplot(data_subset, aes(x = data_subset$bathrooms, y = data_subset$price)) +
  geom_point(color = "blue", size = 3) +
  labs(title = "Scatterplot Bathrooms vs Price",
       x = "Square Footage",
       y = "Price") +
  theme_minimal() + annotate("text",
             x = Inf, y = -Inf,
             label = paste("Pearson's r =", round(cor(x=data_subset$bathrooms, y=data_subset$price, method="pearson", use="complete.obs"), 2)),
             hjust = 1.1, vjust = -0.5,
             size = 4, color = "red")
```

The "Bathrooms vs Price" Scatterplot shows House price for the number of bathrooms in a home. It has a pearson's r of 0.51. Thus, there is a slight linear trend where more bathrooms increase increased house price. Since bathrooms is discrete the plot has price values at certain intervals.
I think its as useful as the "Square Footage vs Price" plot. They both show the linear correlative trend between price and an IV.